---
title: "Algorithmic Attribution"
author: "Trevor Paulsen"
output: html_notebook
---

## Loading the data
Load the Data Feed data, classifications, and friendly names into Spark.

```{r}
# Load the libraries
library(dplyr)
library(sparklyr)
setwd("/data")

# Open a Spark connection
sc = spark_connect(master="local", version="2.2.0", config=sconfig)

# Or if you can connect to a cluster:
sc = spark_connect(master="spark://IP_ADDRESS:7077", spark_home="/[SPARK_HOME_DIRECTORY]/spark/spark-2.2.0-bin-hadoop2.7/", version="2.2.0", config=sconfig)

# Read in the Data Feed Files
data_feed_local = spark_read_csv(
  sc=sc, 
  name="data_feed", 
  path="01-my.data.feed_2018-*.tsv.gz", 
  header=FALSE, 
  delimiter="\t"
)

# Read in my classification file
campaign_class_local = spark_read_csv(
  sc = sc, 
  name = "campaign_class", 
  path = "SC_EXPORT_campaigns_classifications_file.tab", 
  header = TRUE, 
  delimiter = "\t"
)

# Read Lookup File (for "friendly" names)
# This file comes with the Data Feed
col_names = read.csv(
  file="column_headers.tsv", 
  header=FALSE, 
  sep="\t"
)

# Pre-baked preview
t(col_names)
```


## Data Prep
Here, we'll give the Data Feed data some friendly names, and join the classification data to it.

```{r}
# Give data friendly names
data_feed_tbl = data_feed_local %>%
  mutate(
    merged_visitor_ids = paste0(V1,"_",V2),
    fixed_event_list = paste0(",",V5,",")
  ) %>%
  select(
    visitor_id = merged_visitor_ids,
    visit_num = V3,
    hit_time_gmt = V4,
    post_event_list = fixed_event_list,
    post_campaign = V6
  )

data_feed_tbl = data_feed_tbl %>%
  left_join(campaign_class_local, by=c("post_campaign"="Key"))

data_feed_tbl = data_feed_tbl %>%
  mutate(
    mid_campaign = ifelse(post_event_list %regexp% ",20,", Channel, NA),
    conversion = ifelse(post_event_list %regexp% ",1,", 1, 0)
  )

# Prebaked preview:
data_peek1

```


## Attribution Prep
For the next step, we're going to create a new column to isolate individual sequences of events that ended in an order.

```{r}
# Constructing order sequences for all visitors
data_feed_tbl = data_feed_tbl %>%
  group_by(visitor_id) %>%
  arrange(hit_time_gmt) %>%
  mutate(order_seq = ifelse(conversion > 0, 1, NA)) %>%
  mutate(order_seq = lag(cumsum(ifelse(is.na(order_seq), 0, order_seq)))) %>%
  mutate(order_seq = ifelse((row_number() == 1) & (post_event_list %regexp% ",1,"), -1, ifelse(row_number() == 1, 0, order_seq))) %>%
  group_by(visitor_id, order_seq) %>%
  mutate(
    order_seq = ifelse(sum(conversion)>0, order_seq, NA)
  ) %>% ungroup() %>% ungroup()

# Prebaked preview:
data_peek2
```


# Rules-Based Multi-Touch Attribution
Now we'll show you how to do a few position based attribution rules before we do the algorithmic stuff.

```{r}
# Focus only on sequences that actually had an order
attributable_sequences = data_feed_tbl %>%
  filter(!is.na(mid_campaign) & !is.na(order_seq))

# Last touch attribution
last_touch_orders = attributable_sequences %>%
  group_by(visitor_id, order_seq) %>%
  top_n(1,hit_time_gmt) %>%
  select(visitor_id, order_seq, mid_campaign) %>%
  ungroup() %>%
  group_by(mid_campaign) %>%
  summarize(orders = n()) %>% collect()

# First touch attribution
first_touch_orders = attributable_sequences %>%
  group_by(visitor_id, order_seq) %>%
  top_n(1,-hit_time_gmt) %>%
  select(visitor_id, order_seq, mid_campaign) %>%
  ungroup() %>%
  group_by(mid_campaign) %>%
  summarize(orders = n()) %>% collect()

# Linear (even) attribution
linear_orders = attributable_sequences %>%
  group_by(visitor_id, order_seq) %>%
  mutate(
    order_participation = 1/n()
  ) %>%
  ungroup() %>%
  group_by(mid_campaign) %>%
  summarize(orders = sum(order_participation)) %>% collect()
linear_orders$orders = round(linear_orders$orders)

# Time decay (half life) attribution
half_life = 7*24*60*60
half_life_orders = attributable_sequences %>%
  group_by(visitor_id, order_seq) %>%
  mutate(
    order_participation = 0.5^((hit_time_gmt - min(hit_time_gmt))*1/half_life),
    order_participation = order_participation/sum(order_participation)
  ) %>%
  ungroup() %>%
  group_by(mid_campaign) %>%
  summarize(orders = sum(order_participation)) %>% collect()
half_life_orders$orders = round(half_life_orders$orders)

# U shaped (AKA "starter player closer")
u_shaped_orders = attributable_sequences %>%
  group_by(visitor_id, order_seq) %>%
  mutate(
    order_participation = 
      ifelse(n() == 1, 1,
        ifelse(n() == 2, 0.5,
          ifelse(hit_time_gmt == max(hit_time_gmt), 0.4,
            ifelse(hit_time_gmt == min(hit_time_gmt), 0.4,
              0.2/(n()-2)
            )
          )
        )
      )
  ) %>%
  ungroup() %>%
  group_by(mid_campaign) %>%
  summarize(orders = sum(order_participation)) %>% collect()
u_shaped_orders$orders = round(u_shaped_orders$orders)




# Prebaked preview:
p <- plot_ly(model_comparison1, x = ~ model_comparison1$Marketing.Channel, y = ~model_comparison1$Last.Touch, type = 'bar', name = 'Last Touch') %>%
  add_trace(y = ~model_comparison1$First.Touch, name = 'First Touch') %>%
  add_trace(y = ~model_comparison1$Linear, name = 'Linear') %>%
  add_trace(y = ~model_comparison1$Half.Life, name = 'Half Life') %>%
  add_trace(y = ~model_comparison1$U.Shaped, name = 'U-Shaped') %>%
  layout(yaxis = list(title = 'Orders'), barmode = 'group') %>%
  layout(xaxis = list(title = 'Marketing Channel'))
p
model_comparison1

```


## Algorithmic Attribution
Now we'll explore how to do attribution in a smarter way using a cooperative game theory concept known as Shapley Value.
This is more complex, and you can find a more detailed explanation on how to do Shapley Value and more at Trevor's blog here: 
http://datafeedtoolbox.com/attribution-theory-the-two-best-models-for-algorithmic-marketing-attribution-implemented-in-apache-spark-and-r/

```{r}
# Summarizing Order Sequences
seq_summaries = data_feed_tbl %>%
  group_by(visitor_id, order_seq) %>%
  summarize(
    email_touches = max(ifelse(mid_campaign == "Email",1,0)),
    natural_search_touches = max(ifelse(mid_campaign == "Natural_Search",1,0)),
    affiliate_touches = max(ifelse(mid_campaign == "Affiliates",1,0)),
    paid_search_touches = max(ifelse(mid_campaign == "Paid_Search",1,0)),
    display_touches = max(ifelse(mid_campaign == "Display",1,0)),
    social_touches = max(ifelse(mid_campaign == "Social_Media",1,0)),
    conversions = sum(conversion)
  ) %>% ungroup()

conv_rates = seq_summaries %>%
  group_by(email_touches,
           natural_search_touches,
           affiliate_touches,
           paid_search_touches,
           display_touches,
           social_touches) %>%
  summarize(
    conversions = sum(conversions),
    total_sequences = n()
  ) %>% collect()

library(GameTheoryAllocation)

number_of_channels = 6
touch_combos = as.data.frame(coalitions(number_of_channels)$Binary)
names(touch_combos) = c("Email","Natural_Search","Affiliates","Paid_Search","Display","Social_Media")

touch_combo_conv_rate = left_join(touch_combos, conv_rates, 
                                  by = c(
                                    "Email"="email_touches",
                                    "Natural_Search" = "natural_search_touches",
                                    "Affiliates" = "affiliate_touches",
                                    "Paid_Search" = "paid_search_touches",
                                    "Display" = "display_touches",
                                    "Social_Media" = "social_touches"
                                  )
)

touch_combo_conv_rate = touch_combo_conv_rate %>%
  mutate_all(funs(ifelse(is.na(.),0,.))) %>%
  mutate(
    conv_rate = ifelse(total_sequences >0, conversions/total_sequences, 0)
  )


#Building Shapley Values for each combo
shap_vals = as.data.frame(coalitions(number_of_channels)$Binary)
names(shap_vals) = c("Email","Natural_Search","Affiliates","Paid_Search","Display","Social_Media")
coalition_mat = shap_vals
shap_vals[2^number_of_channels,] = Shapley_value(touch_combo_conv_rate$conv_rate, game="profit")

for(i in 2:(2^number_of_channels-1)){
  if(sum(coalition_mat[i,]) == 1){
    shap_vals[i,which(shap_vals[i,]==1)] = touch_combo_conv_rate[i,"conv_rate"]
  }else if(sum(coalition_mat[i,]) > 1){
    if(sum(coalition_mat[i,]) < number_of_channels){
      channels_of_interest = which(coalition_mat[i,] == 1)
      char_func = data.frame(rates = touch_combo_conv_rate[1,"conv_rate"])
      for(j in 2:i){
        if(sum(coalition_mat[j,channels_of_interest])>0 & sum(coalition_mat[j,-channels_of_interest])==0)
          char_func = rbind(char_func,touch_combo_conv_rate[j,"conv_rate"])
      }
      shap_vals[i,channels_of_interest] = Shapley_value(char_func$rates, game="profit")
    } 
  }
}


# Prebaked preview:
p <- plot_ly(model_comparison2, x = ~ model_comparison2$Marketing.Channel, y = ~model_comparison2$Last.Touch, type = 'bar', name = 'Last Touch') %>%
  add_trace(y = ~model_comparison2$First.Touch, name = 'First Touch') %>%
  add_trace(y = ~model_comparison2$Linear, name = 'Linear') %>%
  add_trace(y = ~model_comparison2$Half.Life, name = 'Half Life') %>%
  add_trace(y = ~model_comparison2$U.Shaped, name = 'U-Shaped') %>%
  add_trace(y = ~model_comparison2$Shapley.Value, name = 'Algorithmic') %>%
  layout(yaxis = list(title = 'Orders'), barmode = 'group') %>%
  layout(xaxis = list(title = 'Marketing Channel'))
p
model_comparison2
```








