---
title: "Load Into Workspace"
author: "Trevor Paulsen"
output: html_document
---

## Convert Data Feed Visitor IDs to Hex
The first step is translating the visitor IDs into a format that customer attributes will understand.

```{r}
library(Rmpfr)
library(parallel)
library(RCurl)

# Convert the integer visitor IDs to hex visitor IDs
convert_to_hex = function(visitor_id){
  visid_hi = toupper(as.character(formatMpfr(mpfr(strsplit(visitor_id,"_")[[1]][1], base=10), base=16, drop0trailing=TRUE)))
  visid_lo = toupper(as.character(formatMpfr(mpfr(strsplit(visitor_id,"_")[[1]][2], base=10), base=16, drop0trailing=TRUE)))
  visitor_id_hex = paste0(visid_hi,"-",visid_lo)
  return(visitor_id_hex)
}
n_cores = detectCores() - 1
cl = makeCluster(n_cores, type="FORK")
hex_visitor_ids = parLapply(cl, local_visitor_rollup$visitor_id, convert_to_hex)
stopCluster(cl)
hex_visitor_ids = as.data.frame(unlist(hex_visitor_ids))

# Prebaked Example:
visitor_ids_converted_example

```


## Save off file for Customer Attribute Import
Next, we'll save a file that can be imported to the Customer Attributes feature.

```{r}
# Combine propensity scores with hex values we created
propensity_mapping = data.frame(hex_visitor_ids, propensity_scores)
names(propensity_mapping) = c("aaid", "propensity_scores")

# Write the lookup to a file
write.table(propensity_mapping, file="propensity_mapping.csv", sep=",", row.names = FALSE, quot=FALSE)

# Prebaked Example
propensity_mapping_example
```


## Fire off "setCustomerID" a bunch of times
Now we'll actually load the data into Analytics by spoofing a setCustomerID API call.
I've heard there's a way to do this via batch ingest, but it's not documented...

What's cool about this is you don't even have to have MCID implemented for this to work!

```{r}
# Parallel setCustomerID a zillion times
responses = mclapply(cluster_mapping$aaid, function(aaid){
  org_id = "<YOUR IMS ORG>@AdobeOrg"
  url = paste0("https://dpm.demdex.net/id?",
               "d_visid_ver=1.8.0&",
               "d_fieldgroup=AAM&",
               "d_rtbd=json&",
               "d_ver=2&",
               "d_orgid=", org_id, "&",
               "d_nsid=0&",
               "d_blob=NRX38WO0n5BH8Th-nqAG_A&",
               "d_cid_ic=AVID%01", aaid, "&",
               "d_cid_ic=aaid%01", aaid, "&",
               "ts=1491014875614"
  )
  response = getURL(url = url, ssl.verifyhost = FALSE, ssl.verifypeer = FALSE, followlocation = TRUE)
  return(response)
}, mc.cores=7)

```

